---
title: "Take Home Ex 02: DataVis Makeover"
author: "Roger Chen"
date: "26 Jan 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
number-sections: true
editor: visual
---

# Background

In this take home exercise, we will:

-   select one of the Take-home Exercise 1 prepared by my classmate, [Freddie Jr. Ngo Tan](https://isss608-ay2023-24t2-ftj72.netlify.app/take-home_ex/take-home_ex01/take-home_ex01#insights)

-   critic the submission in terms of clarity and aesthetics,

-   prepare a sketch for the alternative design by using the data visualisation design principles and best practices we had learned in Lesson 1 and 2, and

-   remake the original design by using ggplot2, ggplot2 extensions and tidyverse packages.

## Recap of Take Home Exercise 01

The aim of Take Home Exercise 01 was to use appropriate **Exploratory Data Analysis (EDA)** methods and **ggplot2** functions to reveal:

-   the distribution of Singapore students’ performance in mathematics, reading, and science, and

-   the relationship between these performances with schools, gender and socioeconomic status of the students.

The dataset used was extracted from the "[PISA 2022 database](https://www.oecd.org/pisa/data/2022database/)" which contains the full set of responses from individual students.

# Data Preparation

## Loading R Packages

In this take-home exercise, two R packages will be used. They are:

-   [tidyverse](https://www.tidyverse.org/), and

-   [haven](https://haven.tidyverse.org/).

The code chunk used is as follows:

```{r}
pacman::p_load(tidyverse, haven, knitr, kableExtra, intsvy, ggrepel, patchwork, 
               ggthemes, hrbrthemes, ggdist, ggridges, colorspace, gridExtra)
```

## Importing PISA 2022 Data

The code chunk below uses the [read_sas](https://haven.tidyverse.org/reference/read_sas.html) function from the haven package to import PISA data into R.

```{r}
#| eval: false 

stu_qqq <- read_sas("data/stu_qqq.sas7bdat")
```

We will next filter the observations to those from Singapore, using the [filter](https://dplyr.tidyverse.org/reference/filter.html) function from the dpylr package.

```{r}
#| eval: false 

stu_qqq_SG <- stu_qqq %>%
  filter(CNT == "SGP")
```

We will then save the file as a rds document in the data folder, using [write_rds](https://readr.tidyverse.org/reference/read_rds.html) function. This reduces the file size from 3.9GB to 65.3MB, which makes it easier to push to Github subsequently.

```{r}
#| eval: false 

write_rds(stu_qqq_SG, "data/stu_qqq_SG.rds")
```

We will now import the file back into the R enviroment using the [read_rds](https://readr.tidyverse.org/reference/read_rds.html) function.

```{r}
stu_qqq_SG <- read_rds("data/stu_qqq_SG.rds")
```

::: callout-warning
Remember to:

-   delete the 3.9GB source file from the data folder; and

-   include "#\| eval: false" into the first three code chunks of this section.
:::
