---
title: "Take-Home_Ex04: Rain, Hail or Shine: Unveiling Mysteries of the Sky"
subtitle: "Working Document for Project Work"
author: "Roger Chen"
date: "24 Feb 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
  warning: false
  message: false
number-sections: true
---

# Overview

In this take home exercise, we will

# Data Preparation

## Loading R Packages

In this take home exercise, the following R packages will be used:

The code chunk used is as follows:

```{r}
pacman::p_load(readr, dplyr, purrr, naniar,kableExtra,
               tidyverse, DT, ggplot2, patchwork, 
               ggridges, ggdist, ggstatsplot,
               plotly, ggiraph, DT, ggdist, 
               ungeviz, crosstalk, 
               viridis)
```

## Importing Weather Data

The data sets will be downloaded from [historical daily temperature](https://www.weather.gov.sg/climate-historical-daily/) from [Meteorological Service Singapore](http://www.weather.gov.sg/home/) website,

filter by rain, and temp, load as rds file for faster data wrangling

read rds file

```{r}
#| eval: false
all_data <- read_csv("data/daily_historical.csv")

rain_data <- all_data %>%
  select(station, year, month, day, daily_rainfall_total, highest_30_min_rainfall, highest_60_min_rainfall, highest_120_min_rainfall)

write_rds(rain_data, "data/rain_data.rds")

temp_data <- all_data %>%
  select(station, year, month, day, mean_temperature, maximum_temperature, minimum_temperature)

write_rds(temp_data, "data/temp_data.rds")
```

# Temperature Data EDA

```{r}
temp_data <- read_rds("data/temp_data.rds")

glimpse(temp_data)
```

# Data Wrangling

## Summary Statistics of Data

finding how many missing data

```{r}
sum(is.na(temp_data))
```

dropping those with missing data

```{r}
temp_data <- temp_data %>% drop_na()
```

checking data for missing data again

```{r}
sum(is.na(temp_data))
```

## Aggregating the Mean, Min and Max Temperatures by Month

finding the average, as well as the min and max temperatures by station-month-year

```{r}
monthly_temp_data <- aggregate(mean_temperature ~ year + month, data = temp_data, FUN = mean)
```

```{r}
monthly_mean_temp <- aggregate(mean_temperature ~ station + year + month, data = temp_data, FUN = mean)
```

```{r}
monthly_min_temp <- aggregate(minimum_temperature ~ station + year + month, data = temp_data, FUN = min)
```

```{r}
monthly_max_temp <- aggregate(maximum_temperature ~ station + year + month, data = temp_data, FUN = max)
```

combining the data together

```{r}
merged_temp <- monthly_mean_temp %>%
  left_join(monthly_max_temp, by = c("station", "year", "month")) %>%
  left_join(monthly_min_temp, by = c("station", "year", "month"))

glimpse(merged_temp)
```

Distribution of mean temperature using a ridgeline plot

```{r}

ridge_plot <- ggplot(monthly_temp_data, aes(x = mean_temperature, y=factor(year), fill = stat(x))) +
  geom_density_ridges_gradient(scale =2,rel_min_height = 0.01, gradient_lwd = 1.) +
                       scale_y_discrete(name= NULL)+
                       scale_fill_viridis_c(name = "°C", option = "C") +
                       labs(title = 'temperature profile',
                        subtitle = 'Distribution of daily mean temperature across years',
                        x = "Temperature (°C)",
                        y = "Year") +
                        theme_ridges(font_size = 13, grid = TRUE) +
  theme(plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 10),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8, angle = 360))

ridge_plot

```

Distribution of mean temperature using a boxplot

```{r}
box_plot <- ggplot(monthly_temp_data, 
                   aes(y= mean_temperature,
                       x = as.factor(year))) +
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 60))

box_plot
```

Distribution of mean temperature using a boxplot, by stations

```{r}
#| fig-width: 15
#| fig-height: 12

box_plot_m <- ggplot(merged_temp, 
                   aes(y= mean_temperature,
                       x = as.factor(year))) +
  geom_boxplot()+
  facet_wrap(~ station) +
  theme(axis.text.x = element_text(angle = 60))

box_plot_m
```

```{r}
violin_plot <- plot_ly(data = monthly_temp_data, x = ~year, y = ~mean_temperature, type = 'violin', 
               split = ~year, 
               box = list(visible = T),
               meanline = list(visible = T)) %>% # Adds a box plot inside the violin for summary statistics
  layout(title = list(text = "<b>Distribution of Mean Temperatures Across Years</b>", 
                                                   font = list(family = "Arial", size = 14),
                                                   xanchor = "center", xref = "paper",
                                                   yanchor = "middle", yref = "paper"),
                      yaxis = list(title = list(text = "Temperature (°C)", 
                                                font = list(family = "Arial", size = 12)) ),
                      xaxis = list(title = list(text = "Year", 
                                                font = list(family = "Arial", size = 12)) ))

violin_plot
```

# Rain Data EDA

```{r}
rain_data <- read_rds("data/rain_data.rds")

glimpse(rain_data)
```

## Summary Statistics of Data

finding how many missing data

```{r}
sum(is.na(rain_data))
```

dropping those with missing data

```{r}
rain_data <- rain_data %>% drop_na()

```

checking data for missing data again

```{r}
sum(is.na(rain_data))

```

# Static Data Visualisation

::: panel-tabset
## Plot

```{r, fig.width = 15, fig.height = 8}
#| column: page
#| echo: false


          

```

## Code Chunk

```{r}
#| eval: false



```
:::

::: callout-note
:::

# Interactive Data Visualisation

::: panel-tabset
## Plot

```{r, fig.height=12}
#| column: page
#| echo: false


```

## Code Chunk

```{r}
#| eval: false


```
:::

::: callout-note
:::

::: panel-tabset
## Plot

```{r}
#| echo: false

      
```

## Code Chunk

```{r}
#| eval: false




```
:::

::: callout-note
:::

# Conclusion
