---
title: "Take-Home_Ex04: Rain, Hail or Shine: Unveiling Mysteries of the Sky"
subtitle: "Working Document for Project Work"
author: "Roger Chen"
date: "24 Feb 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
  warning: false
  message: false
number-sections: true
---

# Overview

In this take home exercise, we will

# Data Preparation

## Loading R Packages

In this take home exercise, the following R packages will be used:

The code chunk used is as follows:

```{r}
pacman::p_load(readr, dplyr, purrr, naniar,kableExtra,
               tidyverse, DT, ggplot2, patchwork, 
               ggridges, ggdist, ggstatsplot,
               plotly, ggiraph, DT, ggdist, 
               ungeviz, crosstalk, 
               viridis)
```

## Importing Weather Data

The data sets will be downloaded from [historical daily temperature](https://www.weather.gov.sg/climate-historical-daily/) from [Meteorological Service Singapore](http://www.weather.gov.sg/home/) website,

filter by rain, and temp, load as rds file for faster data wrangling

read rds file

```{r}
#| eval: false
all_data <- read_csv("data/daily_historical.csv")

rain_data <- all_data %>%
  select(station, year, month, day, daily_rainfall_total, highest_30_min_rainfall, highest_60_min_rainfall, highest_120_min_rainfall)

write_rds(rain_data, "data/rain_data.rds")

temp_data <- all_data %>%
  select(station, year, month, day, mean_temperature, maximum_temperature, minimum_temperature)

write_rds(temp_data, "data/temp_data.rds")
```

# Rain Data EDA

```{r}
rain_data <- read_rds("data/rain_data.rds")

glimpse(rain_data)
```

## Summary Statistics of Data

finding how many missing data

```{r}
sum(is.na(rain_data))
```

dropping those with missing data

```{r}
rain_data <- rain_data %>% drop_na()

```

checking data for missing data again

```{r}
sum(is.na(rain_data))

```

# Static Data Visualisation

::: panel-tabset
## Plot

```{r, fig.width = 15, fig.height = 8}
#| column: page
#| echo: false


          

```

## Code Chunk

```{r}
#| eval: false



```
:::

::: callout-note
:::

# Interactive Data Visualisation

::: panel-tabset
## Plot

```{r, fig.height=12}
#| column: page
#| echo: false


```

## Code Chunk

```{r}
#| eval: false


```
:::

::: callout-note
:::

::: panel-tabset
## Plot

```{r}
#| echo: false

      
```

## Code Chunk

```{r}
#| eval: false




```
:::

::: callout-note
:::

# Conclusion
