---
title: "Take-Home_Ex04: Rain, Hail or Shine: Unveiling Mysteries of the Sky"
subtitle: "Working Document for Project Work"
author: "Roger Chen"
date: "24 Feb 2024"
date-modified: "last-modified"
execute: 
  warning: false
  freeze: true
number-sections: true
---

# Overview

In this take home exercise, we will

# Data Preparation

## Loading R Packages

In this take home exercise, the following R packages will be used:

-   [tidyverse](https://www.tidyverse.org/),

-   [ggplot2](https://ggplot2.tidyverse.org/),

-   [gganimate](https://gganimate.com/),

-   [plotly](https://plotly.com/r/),

-   [ggiraph](https://davidgohel.github.io/ggiraph/),

-   [DT](https://rstudio.github.io/DT/).

The code chunk used is as follows:

```{r}
pacman::p_load(tidyverse, ggplot2, gganimate, plotly, ggiraph, DT)
```

## Importing Temperature Data

***Changi*** will be selected for the weather station, and ***temperature*** chosen as the factor to be analysed. The data sets will be downloaded from [historical daily temperature](https://www.weather.gov.sg/climate-historical-daily/) from [Meteorological Service Singapore](http://www.weather.gov.sg/home/) website,

Lastly, using the code chunk below, we will combine the five datasets into a single document, and save it as a new dataset.

Next, we will call the dataset "***combinedTemp***" into the environment.

```{r}
data <- read_csv("data/daily_historical.csv")
```

## Summary Statistics of Data

::: panel-tabset
### First 5 Rows

Using [**DT**](https://rstudio.github.io/DT/), we will display the dataset as a interactive table on html page.

```{r}
DT::datatable(data, class = "display compact", style = "bootstrap5")
```

::: callout-note
The data table seemed to be in order.
:::

### Dataset Structure

```{r}
str(data)
```

::: callout-note
The data types are all correct.
:::

### Data Health

Checking for missing values

```{r}
sum(is.na(data))
```

::: callout-note
No missing values were found.
:::
:::

# Static Data Visualisation

::: panel-tabset
## Plot

```{r, fig.width = 15, fig.height = 8}
#| column: page
#| echo: false


          

```

## Code Chunk

```{r}
#| eval: false



```
:::

::: callout-note
:::

# Interactive Data Visualisation

::: panel-tabset
## Plot

```{r, fig.height=12}
#| column: page
#| echo: false


```

## Code Chunk

```{r}
#| eval: false


```
:::

::: callout-note
:::

::: panel-tabset
## Plot

```{r}
#| echo: false

      
```

## Code Chunk

```{r}
#| eval: false




```
:::

::: callout-note
:::

# Conclusion
